---
title: 'Part 1: Binary Logistic Regression'
author: "Katelin Seeto"
date: "1/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
library(GGally) #make pairs plots
library(broom) # to get tidy model outputs in df's
library(jtools) # make quick graphs of model outputs
```

### Create pairs plot with GGally

Notes:  
* `ggpairs()` creates plots based on the variable you select (in this case we chose species) that shows multivariate correlations, distributions, and relationships both within and across variables
* `fct_drop()` will drop factor levels where there's no observations for those factor levels

```{r}
penguins %>% 
  select(species, bill_length_mm:sex) %>% 
  ggpairs(aes(color = species))
```

Create a subset with only Adelie and Chinstrap penguins. Remember that species are listed stored as factor with levels

```{r}
adelie_chinstrap <- penguins %>% 
  filter(species %in% c("Adelie", "Chinstrap")) %>% 
# Since species was a factor, even if you filter out all observations for Gentoo, it still appears as a factor level even if it doesn't appear in your df, therefore must use fct_drop
  mutate(species = fct_drop(species))
```

Dataviz of Adelie and Chinstrap: 

```{r}
ggplot(data = adelie_chinstrap, aes(x = body_mass_g, y = flipper_length_mm))+ 
  geom_point(aes(color = sex))+ 
  facet_wrap(~species)
```
Binary logistic regression 






